Weak interspecific interactions in a sagebrush steppe: evidence from observations, models, and experiments

### 2017-05-12 removals/manuscript/removal_ms.tex
SPE edits

Trends in cover by species and removal treatment were variable. For example, grass removal had little effect on the cover of \textit{A. tripartita} 
(Fig. \ref{fig:CoverTrends}), but shrub removal did appear to cause an increase in the cover of \textit{P. spicata}. 
However, \new{when we tested the effects of removal on cover by analyzing year-to-year changes in (log) cover, a measure of population 
growth rate, none of the removal treatment effects were significantly different from zero (Table \ref{table:coefficients}). For \textit{P. spicata}, 
the 95\% confidence interval only just contained zero (95\% CI: -0.01, 0.38), providing marginal evidence for a positive effect of removal on population growth. 
Fitting with weighted least squares to address heteroscedasticity in residuals indicated even less evidence for a positive 
response of  \textit{P. spicata} to removal.} 
\comment{SPE: I think you need to explain how did ``analyzing year-to-year changes in (log) cover'' -- what was the response, what were the
covariates? I think reviewers will also want to know what was done. And without that information, 
it's not clear what the residuals would be, or what weighted least squares would entail. } 
However, these treatment comparisons do not directly address our questions about whether 
responses to removal were adequately predicted by models fit to historical, observational data.

@@

Our vital rate models take into account the effects of local species composition and should therefore be able to explain the effect of removals on survival, growth and recruitment without any additional parameters. 
If the removal treatment effects that we added to these models are statistically significant, it would indicate an effect of removals either larger or smaller than the effect predicted by a model based on data 
from the historical control plots. 
\comment{SPE: is the baseline model just based on the historical control plots data, or also on the modern data from control plots?} 
For the survival models, the 95\% confidence intervals on the removal effects overlapped zero for all four species (Fig. \ref{fig:VitalRateTest}; 
Tables \ref{ARTRsurvival} to \ref{PSSPsurvival}), meaning that the survival probabilities of plants in removal plots can be predicted based on data 
from control plots. For the growth models, the 95\% confidence intervals on the removal effects were positive for \textit{P. secunda}  and \textit{P. spicata},
 meaning that individuals of these species grew more rapidly in removal plots than controls even after accounting for neighborhood interactions 
 (Fig. \ref{fig:VitalRateTest}; Tables \ref{ARTRgrowth} to \ref{PSSPgrowth}). In contrast, \textit{Poa secunda} had lower recruitment in removal 
 than controls plots (upper limit of 95\% CI $<0$; Fig. \ref{fig:VitalRateTest}; Table \ref{table:recruitment}). For the other species, 
 we could not detect significant removal treatment effects on recruitment.

@@

\begin{table}
\centering
\caption{Summary of fixed effects for the \textit{P. secunda} model with treatment*year effects (the ``trtYears*" coefficients). See Table \ref{ARTRgrowth} for an explanation of other coefficient names and column header.} 
\label{table:POSEgrowth-trtYears}
\begin{tabular}{rrrrrrrr}
  \hline
 & mean & sd & 0.025quant & 0.5quant & 0.975quant & mode & kld \\ 
  \hline
(Intercept) & 0.4963 & 0.0616 & 0.3752 & 0.4963 & 0.6172 & 0.4964 & 0.0000 \\ 
  trtYears1 & 0.0798 & 0.1272 & -0.1699 & 0.0798 & 0.3293 & 0.0798 & 0.0000 \\ 
  trtYears2 & 0.2425 & 0.1350 & -0.0225 & 0.2425 & 0.5073 & 0.2425 & 0.0000 \\ 
  trtYears3 & 0.2512 & 0.1554 & -0.0539 & 0.2512 & 0.5561 & 0.2512 & 0.0000 \\ 
  trtYears4 & 0.2988 & 0.1693 & -0.0337 & 0.2988 & 0.6310 & 0.2988 & 0.0000 \\ 
  trtYears5 & 0.4787 & 0.1845 & 0.1164 & 0.4787 & 0.8407 & 0.4787 & 0.0000 \\ 
  logarea.t0 & 0.6731 & 0.0234 & 0.6270 & 0.6732 & 0.7191 & 0.6732 & 0.0000 \\ 
  W.ARTR & -0.0003 & 0.0009 & -0.0019 & -0.0003 & 0.0014 & -0.0003 & 0.0000 \\ 
  W.HECO & 0.0035 & 0.0064 & -0.0090 & 0.0035 & 0.0160 & 0.0035 & 0.0000 \\ 
  W.POSE & -0.2624 & 0.0394 & -0.3398 & -0.2624 & -0.1851 & -0.2624 & 0.0000 \\ 
  W.PSSP & -0.0080 & 0.0057 & -0.0192 & -0.0080 & 0.0032 & -0.0080 & 0.0000 \\ 
  W.allcov & -0.0004 & 0.0012 & -0.0027 & -0.0004 & 0.0020 & -0.0004 & 0.0000 \\ 
  W.allpts & -0.0267 & 0.0236 & -0.0730 & -0.0267 & 0.0196 & -0.0267 & 0.0000 \\ 
  inARTR & -0.0563 & 0.1169 & -0.2859 & -0.0563 & 0.1730 & -0.0563 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}
\centering
\caption{Summary of fixed effects for the \textit{P. spicata} model with treatment*year effects (the ``trtYears*" coefficients). See Table \ref{ARTRgrowth} for an explanation of other coefficient names and column headers.} 
\label{table:PSSPgrowth-trtYears}
\begin{tabular}{rrrrrrrr}
  \hline
 & mean & sd & 0.025quant & 0.5quant & 0.975quant & mode & kld \\ 
  \hline
(Intercept) & 0.4122 & 0.0626 & 0.2892 & 0.4120 & 0.5364 & 0.4115 & 0.0000 \\ 
  trtYears1 & 0.3256 & 0.1440 & 0.0428 & 0.3256 & 0.6081 & 0.3257 & 0.0000 \\ 
  trtYears2 & -0.0353 & 0.1467 & -0.3235 & -0.0353 & 0.2525 & -0.0353 & 0.0000 \\ 
  trtYears3 & 0.4829 & 0.1586 & 0.1714 & 0.4829 & 0.7940 & 0.4829 & 0.0000 \\ 
  trtYears4 & 0.5059 & 0.1602 & 0.1915 & 0.5059 & 0.8202 & 0.5059 & 0.0000 \\ 
  trtYears5 & -0.0588 & 0.1682 & -0.3890 & -0.0588 & 0.2712 & -0.0588 & 0.0000 \\ 
  logarea.t0 & 0.8266 & 0.0173 & 0.7922 & 0.8267 & 0.8607 & 0.8268 & 0.0000 \\ 
  W.ARTR & -0.0030 & 0.0008 & -0.0044 & -0.0030 & -0.0015 & -0.0030 & 0.0000 \\ 
  W.HECO & -0.0139 & 0.0068 & -0.0272 & -0.0139 & -0.0005 & -0.0139 & 0.0000 \\ 
  W.POSE & -0.0064 & 0.0127 & -0.0313 & -0.0064 & 0.0185 & -0.0064 & 0.0000 \\ 
  W.PSSP & -0.1362 & 0.0126 & -0.1609 & -0.1362 & -0.1115 & -0.1362 & 0.0000 \\ 
  W.allcov & -0.0040 & 0.0010 & -0.0061 & -0.0040 & -0.0020 & -0.0040 & 0.0000 \\ 
  W.allpts & -0.0579 & 0.0224 & -0.1020 & -0.0579 & -0.0139 & -0.0579 & 0.0000 \\ 
  inARTR & -0.1235 & 0.1104 & -0.3403 & -0.1235 & 0.0931 & -0.1235 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}

@@

>>>>

Trends in cover by species and removal treatment were variable. For example, grass removal had little effect on the cover of \textit{A. tripartita} 
(Fig. \ref{fig:CoverTrends}), but shrub removal did appear to cause an increase in the cover of \textit{P. spicata}. 
However, \new{when we tested the effects of removal on cover by analyzing year-to-year changes in (log) cover, a measure of population 
growth rate, none of the removal treatment effects were significantly different from zero (Table \ref{table:coefficients}). For \textit{P. spicata}, 
the 95\% confidence interval only just contained zero (95\% CI: -0.01, 0.38), providing marginal evidence for a positive effect of removal on population growth. 
Fitting with weighted least squares to address heteroscedasticity in residuals indicated even less evidence for a positive 
response of  \textit{P. spicata} to removal.} 
\comment{SPE: I think you need to explain how did ``analyzing year-to-year changes in (log) cover'' -- what was the response, what were the
covariates? I think reviewers will also want to know what was done. And without that information, 
it's not clear what the residuals would be, or what weighted least squares would entail. } 
However, these treatment comparisons do not directly address our questions about whether 
responses to removal were adequately predicted by models fit to historical, observational data. \comment{SPE: is it ``historical, observational data'' or was the ``baseline'' model
fitted to observational data both historical and modern?}

@@

Our vital rate models take into account the effects of local species composition and should therefore be able 
to explain the effect of removals on survival, growth and recruitment without any additional parameters. 
\new{SPE: As in our previous studies using only historical data, the neighborhood competition coefficients in the 
fitted models imply that interspecific competition is much weaker than intraspecific competition, in many
cases by 1-2 orders of magnitude (Table \ref{table:alphas}).}  
If the removal treatment effects that we added to these models are statistically significant, it would indicate an effect of 
removals either larger or smaller than the effect predicted by a model based on data from the historical control plots. 
\comment{SPE: is the baseline model just based on the historical data, or also on the modern data from control plots? 
For me, ``historical control plots'' is ambiguous on this point.} 
For the survival models, the 95\% confidence intervals on the removal effects overlapped zero for all four species (Fig. \ref{fig:VitalRateTest}; 
Tables \ref{ARTRsurvival} to \ref{PSSPsurvival}), meaning that the survival probabilities of plants in removal plots can be predicted based on data 
from control plots. For the growth models, the 95\% confidence intervals on the removal effects were positive for \textit{P. secunda}  and \textit{P. spicata},
 meaning that individuals of these species grew more rapidly in removal plots than controls even after accounting for neighborhood interactions 
 (Fig. \ref{fig:VitalRateTest}; Tables \ref{ARTRgrowth} to \ref{PSSPgrowth}). In contrast, \textit{Poa secunda} had lower recruitment in removal 
 than controls plots (upper limit of 95\% CI $<0$; Fig. \ref{fig:VitalRateTest}; Table \ref{table:recruitment}). For the other species, 
 we could not detect significant removal treatment effects on recruitment.

@@

\begin{table}
\centering
\caption{Summary of fixed effects for the \textit{P. secunda} growth model with treatment*year effects (the ``trtYears*" coefficients). See Table \ref{ARTRgrowth} for an explanation of other coefficient names and column header.} 
\label{table:POSEgrowth-trtYears}
\begin{tabular}{rrrrrrrr}
  \hline
 & mean & sd & 0.025quant & 0.5quant & 0.975quant & mode & kld \\ 
  \hline
(Intercept) & 0.4963 & 0.0616 & 0.3752 & 0.4963 & 0.6172 & 0.4964 & 0.0000 \\ 
  trtYears1 & 0.0798 & 0.1272 & -0.1699 & 0.0798 & 0.3293 & 0.0798 & 0.0000 \\ 
  trtYears2 & 0.2425 & 0.1350 & -0.0225 & 0.2425 & 0.5073 & 0.2425 & 0.0000 \\ 
  trtYears3 & 0.2512 & 0.1554 & -0.0539 & 0.2512 & 0.5561 & 0.2512 & 0.0000 \\ 
  trtYears4 & 0.2988 & 0.1693 & -0.0337 & 0.2988 & 0.6310 & 0.2988 & 0.0000 \\ 
  trtYears5 & 0.4787 & 0.1845 & 0.1164 & 0.4787 & 0.8407 & 0.4787 & 0.0000 \\ 
  logarea.t0 & 0.6731 & 0.0234 & 0.6270 & 0.6732 & 0.7191 & 0.6732 & 0.0000 \\ 
  W.ARTR & -0.0003 & 0.0009 & -0.0019 & -0.0003 & 0.0014 & -0.0003 & 0.0000 \\ 
  W.HECO & 0.0035 & 0.0064 & -0.0090 & 0.0035 & 0.0160 & 0.0035 & 0.0000 \\ 
  W.POSE & -0.2624 & 0.0394 & -0.3398 & -0.2624 & -0.1851 & -0.2624 & 0.0000 \\ 
  W.PSSP & -0.0080 & 0.0057 & -0.0192 & -0.0080 & 0.0032 & -0.0080 & 0.0000 \\ 
  W.allcov & -0.0004 & 0.0012 & -0.0027 & -0.0004 & 0.0020 & -0.0004 & 0.0000 \\ 
  W.allpts & -0.0267 & 0.0236 & -0.0730 & -0.0267 & 0.0196 & -0.0267 & 0.0000 \\ 
  inARTR & -0.0563 & 0.1169 & -0.2859 & -0.0563 & 0.1730 & -0.0563 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}
\centering
\caption{Summary of fixed effects for the \textit{P. spicata} growth model with treatment*year effects (the ``trtYears*" coefficients). See Table \ref{ARTRgrowth} for an explanation of other coefficient names and column headers.} 
\label{table:PSSPgrowth-trtYears}
\begin{tabular}{rrrrrrrr}
  \hline
 & mean & sd & 0.025quant & 0.5quant & 0.975quant & mode & kld \\ 
  \hline
(Intercept) & 0.4122 & 0.0626 & 0.2892 & 0.4120 & 0.5364 & 0.4115 & 0.0000 \\ 
  trtYears1 & 0.3256 & 0.1440 & 0.0428 & 0.3256 & 0.6081 & 0.3257 & 0.0000 \\ 
  trtYears2 & -0.0353 & 0.1467 & -0.3235 & -0.0353 & 0.2525 & -0.0353 & 0.0000 \\ 
  trtYears3 & 0.4829 & 0.1586 & 0.1714 & 0.4829 & 0.7940 & 0.4829 & 0.0000 \\ 
  trtYears4 & 0.5059 & 0.1602 & 0.1915 & 0.5059 & 0.8202 & 0.5059 & 0.0000 \\ 
  trtYears5 & -0.0588 & 0.1682 & -0.3890 & -0.0588 & 0.2712 & -0.0588 & 0.0000 \\ 
  logarea.t0 & 0.8266 & 0.0173 & 0.7922 & 0.8267 & 0.8607 & 0.8268 & 0.0000 \\ 
  W.ARTR & -0.0030 & 0.0008 & -0.0044 & -0.0030 & -0.0015 & -0.0030 & 0.0000 \\ 
  W.HECO & -0.0139 & 0.0068 & -0.0272 & -0.0139 & -0.0005 & -0.0139 & 0.0000 \\ 
  W.POSE & -0.0064 & 0.0127 & -0.0313 & -0.0064 & 0.0185 & -0.0064 & 0.0000 \\ 
  W.PSSP & -0.1362 & 0.0126 & -0.1609 & -0.1362 & -0.1115 & -0.1362 & 0.0000 \\ 
  W.allcov & -0.0040 & 0.0010 & -0.0061 & -0.0040 & -0.0020 & -0.0040 & 0.0000 \\ 
  W.allpts & -0.0579 & 0.0224 & -0.1020 & -0.0579 & -0.0139 & -0.0579 & 0.0000 \\ 
  inARTR & -0.1235 & 0.1104 & -0.3403 & -0.1235 & 0.0931 & -0.1235 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}

@@

\newpage 
\begin{table}[ht]
\centering
\caption{Summary of competition coefficients in the demographic models. These are the ``W.abcd'' fixed effects coefficients from
the Tables in this section summarizing the survival, growth and recruitment models for each species.  
In each $4 \times 4$ array of coefficients, the entry in row $i$, column $j$ is the 
coefficient for the effect of species $j$ cover on species $i$. Intraspecific competition is weak relative to intraspecific competition if each
species affects itself more than others, i.e. if the largest (in magnitude) element in each column is the element on the diagonal of the array.
The one exception to this pattern is the effect of ARTR on recruitment by POSE, but the nonconforming
coefficient ($\omega$[3,1] = -0.6589 in Table \ref{table:recruitment}) has low imprecision (sd = 0.3342, Table \ref{table:recruitment}.} 
\label{table:alphas}
\begin{tabular}{lrrrr}
       & ARTR &  HECO &  POSE &  PSSP \\
\hline 
\underline{Survival} & & & & \\        
ARTR   & -1.5214  & -0.0694  & -0.0049 &  0.0281\\
HECO   & -0.0064  &  -0.6647 &  0.0732 &  0.0190\\
POSE  &   0.0001&  -0.0153 &  -1.2632&   0.0249\\
PSSP  &  0.0101 &  0.0046 & 0.0245 &  -0.4430\\
\underline{Growth} & & & & \\ 
ARTR   & -0.0364 &  0.0033& -0.0573&  0.0062\\
HECO   &  -0.0031 & -0.0479& 0.0242 & -0.0151\\
POSE  &  -0.0002& 0.0039&  -0.2641&   -0.0073\\
PSSP  & -0.0025 & -0.0118 & -0.0067 & -0.1480\\
\underline{Recruitment} & & & & \\ 
ARTR  & -0.5881 & 0.0635 & 0.0220 & 0.1164 \\
HECO   &  -0.4800 & -1.7368& 0.0388 &  -0.3682\\
POSE  &  -0.6589& -0.1021& -1.8943 &   -0.1953\\
PSSP  & -0.1867 &-0.3746  & 0.1131& -1.7379\\
\hline
\end{tabular}
\end{table}
